---
title: "Lab 4"
author: "Rachel Carnes and Britnie Smith" 
output: word_document
date: '2022-04-09'
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  
n <- 100 
sig <- 1 
b <- c(1.5,-2,-0.5,1) # beta coefficients 
set.seed(10) 
x <- sort(round(runif(n,-2,2),2)) # predictor x 
eps <- rnorm(n,0,sig) # random noise 
ones <- vector(length=n)+1 
x2 <- x^2 
x3 <- x^3 
xmat <- cbind(ones, x, x2, x3) 
head(xmat) 
fx <- xmat%*%b # true nonlinear relationship is cubic  
y <- fx + eps  
data <- data.frame(x,y) # this is the sample data 
plot(x,fx,type='l',lwd=2,ylim=c(-5,5),  
     main='True relationship between y and x with generated sample') 
points(x,y,pch=19)

?runif
View(data)

```

1.  Write out the true functional relationship that is displayed in this
    plot. That is, from the code, identify the coefficients ð›½0,ð›½1,ð›½2,
    and ð›½3 in the generating model ð‘¦= ð›½0 + ð›½1ð‘¥+ ð›½2ð‘¥2 + ð›½3ð‘¥3 +ðœ€.

y =1.5 -2*x -0.5*x2 + 1\*x3 +ðœ€ is the model.

```{r}

```

2.  Use the regsubsets() function to perform best subset selection in
    order to choose the best model from among the predictors x, x 2, x
    3, x 4, ..., x 10.

```{r}

x4 <- x^4
x5 <- x^5
x6 <- x^6
x7 <- x^7
x8 <- x^8
x9 <- x^9
x10 <- x^10
b <- c(1.5,-2,-0.5,1)

xmat2 <- cbind(ones, x, x2, x3, x4, x5, x6, x7, x8, x9, x10) 
head(xmat2) 

library(leaps)

regfit_full <- regsubsets(y ~ x + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10, data = data)

summary (regfit_full)

res.sum <- summary(regfit_full)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)


par(mfrow=c(1,1))
plot(regfit_full, scale = "r2")
plot(regfit_full, scale = "adjr2")
plot(regfit_full, scale = "Cp")
plot(regfit_full, scale = "bic")

dataplot <- lm(y ~ x + x2 + x3, data= data)
dataplot

dataplot2 <- lm(y ~ x + x2 + x3 + x4, data= data)
dataplot2

```

a.  What is the best model obtained according to the Cp criterion?
    Provide the coefficient table of estimates.

Cp is saying the best model has four variables are they are x, x2, x3,
and x4.

b.  What is the best model obtained according to the BIC? Provide the
    coefficient table of estimates.

BIC is saying the best model has three variables which are x, x1 and x3.

c.  What is the best model obtained according to the adjusted R2?
    Provide the coefficient table of estimates.

Adj R2 is saying the best model has four variables are they are x, x2,
x3, and x4.

d.  Plot each of the best fits on the same plot with the true curve and
    the data. Comment on your plot and the estimated fits.

???????????????WHAT???????????????????

```{r}

```

3.  Repeat (2) using forward selection and then backwards selection.
    Compare the results and briefly summarize your findings.

```{r}

regfit.fwd <- regsubsets(y ~ x + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10, data = data, nvmax = 19, method = "forward")
summary(regfit.fwd)

res.sum <- summary(regfit.fwd)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)

par(mfrow=c(1,1))
plot(regfit.fwd, scale = "r2")
plot(regfit.fwd, scale = "adjr2")
plot(regfit.fwd, scale = "Cp")
plot(regfit.fwd, scale = "bic")

plot (res.sum$rss , xlab = " Number of Variables ",
ylab = " RSS ", type = "l")

regfit.bwd <- regsubsets(y ~ x + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10, data = data,nvmax = 19, method = "backward")
summary(regfit.bwd)

res.sum2 <- summary(regfit.bwd)
data.frame(
  Adj.R2 = which.max(res.sum2$adjr2),
  CP = which.min(res.sum2$cp),
  BIC = which.min(res.sum2$bic)
)

par(mfrow=c(1,1))
plot(regfit.bwd, scale = "r2")
plot(regfit.bwd, scale = "adjr2")
plot(regfit.bwd, scale = "Cp")
plot(regfit.bwd, scale = "bic")

plot (res.sum2$rss , xlab = " Number of Variables ",
ylab = " RSS ", type = "l")

```

4.  Fit a lasso model using the predictors x, x 2, x 3, x 4, ..., x 10.
    Use CV to select the optimal tuning parameter, ï¬. What value of ï¬
    did you use? What is the resulting fitted model?\

```{r}
#what is lambda and what does it do. How is it calculated and how is it used?


x4 <- x^4
x5 <- x^5
x6 <- x^6
x7 <- x^7
x8 <- x^8
x9 <- x^9
x10 <- x^10
b <- c(1.5,-2,-0.5,1)
dataplot <- lm(y ~ x + x2 + x3, data= data)
dataplot

dataplot2 <- lm(y ~ x + x2 + x3 + x4, data= data)
dataplot2

xmat2 <- cbind(ones, x, x2, x3, x4, x5, x6, x7, x8, x9, x10) 
head(xmat2)
lasso1 <- glmnet(xmat2, y, alpha = 1, lambda = 0.005529133) # alpha = 1 corresponds to lasso
cbind(coef(dataplot),coef(lasso1))


grid <- 10^seq(10,-2,length=100) # grid of lambda values goes from 10^10 down to 10^-2 = 0.01
grid  
lasso.mod <- glmnet(xmat2, y, alpha = 1, lambda = grid) 
dim(coef(lasso.mod)) # 20 coefficients by 100 lambda values

set.seed(1)
cv.lasso.mod <- cv.glmnet(xmat2, y, alpha = 1)
plot(cv.lasso.mod)
cv.lasso.mod$lambda.min

lassop <- predict(lasso.mod, s=cv.lasso.mod$lambda.min, type="coefficients" )

lassop


```

5.  Use a smoothing spline (sm.spline()) from the pspline package to
    estimate the nonlinear relationship. Use degrees of freedom set to
    df = 2, 3, 4, 5, 15 and the optimal df chosen by CV. Give the
    optimal df. Plot each of the estimated fits on the same plot with
    the true curve and the data. Comment on your plot and the estimated
    fits.\

```{r}
library(pspline)
?sm.spline

pspline2 <- sm.spline(y,x,spar=150)
pspline2

plot(y,x,pch=19,xlab="y")
lines(pspline,col=5,lty=4,lwd=3)
pspline2 <- sm.spline(age,strontium.ratio,df=2)
pspline3 <- sm.spline(age,strontium.ratio,df=3)
pspline4 <- sm.spline(age,strontium.ratio,df=4)
pspline5 <- sm.spline(age,strontium.ratio,df=5)
pspline15 <- sm.spline(age,strontium.ratio,df=15)


# selecting the smoothing parameter using cross-validation
psplinecv <- sm.spline(age,strontium.ratio,cv=T)
psplinecv
lines(psplinecv,col=1,lty=1,lwd=2)
legtxt <- c("df = 50","df = 25", "df = 5", "df = 3","df CV")
legend("bottomleft",legtxt,lty=c(2,3,4,5,1),col=c(2,3,4,5,1),lwd=2)

```

6.  Which of all estimates examined (out of the 15 ones examined in (2)
    -- (5)) provides the best fit to the true relationship, fx? Give
    support for your response (provide a quantitative measure).

```{r}

```

7.  Is it possible to use principal components regression (or partial
    least squares) to estimate the nonlinear relationship with the
    predictors x, x 2, x 3, x 4, ..., x 10? Explain why or why not.\
    ???

```{r}



```

8.  Alter one aspect of the simulation setting (such as the true
    relationship fx, the sample size, or the amount of residual
    variation in the generated data in sig) and revisit one of the
    methods (best subsets, forward selection, backward selection, lasso,
    ridge regression, penalized splines) to see how the method performs
    at estimating fx. Plot the true fx. and your estimate on the same
    plot. Give a measure of performance. Submit code for this question.\

We are looking for the smallest RSS because that shows the best fit for
fx.

```{r}

```

Note that a true simulation study iterates through many randomly
generated data sets under a given scenario to examine properties of the
method and the distribution of the estimate.

```{r}

```
